When do the effects of interaction happen? Interacting and disturbed galaxies in the COSMOS field

### 2024-01-22 paper-source/main.tex
Updated code based on Brookes comments. Preparing to update to send to collaborators.

% Abstract of the paper
\begin{abstract}
   Galaxy interaction and merging have fundamental impacts on galactic evolution. From increases in star formation, to morphology, to activation of nuclear activity, galaxy interaction is thought to play a role. However, at what point in the interaction do these effects occur? In this work, we break down a large sample of identified interacting galaxies and split them into four distinct stages based on their interaction morphology. Each stage captures a different part of the dynamical timescale of interaction; from the systems approaching each other to actually merging. We cross match this large sample with the COSMOS field to provide ancillary data and each galaxys' underlying parameters. We then investigate the relation between stage, mass, star formation, and nuclear activation. Using Kolmogorov-Smirnov tests, we find that the distribution of star formation rates rapidly changes with stage for an equivalent mass distribution. We find a rapidly shrinking red sequence with stage, and a large increase in the fraction of starburst galaxies at the point of coalescence. We also find that the fraction of active galactic nuclei is static through most of the interaction, until it dramatically increases at coalescence. We discuss these results, and compare them with the more prevalent stage diagnostic of projected separation between galaxy pairs. We find interesting behaviour that is morphology and stage dependent in our samples star formation rates which is not captured by only using the projected separation. We show that morphological stage and, more specifically, the point in time of the underlying dynamical timescale an interaction occupies is as important to consider as its projected separation. We conclude this work with discussions on how this could be applied to larger, more complete datasets of interacting galaxies with available ancillary data.
\end{abstract}

@@

\section{Introduction}\label{introduction}
\noindent Galaxy interaction and merging is of fundamental importance to our current theories of galaxy formation and evolution. Our current cosmological theory, $\Lambda$CDM dictates that structure is expected to form and assemble hierarchically \citep{1978MNRAS.183..341W, 2001MNRAS.328..726S, 2013MNRAS.436.1765M}. Thus, understanding the underlying processes that occur in galaxy interaction and merging are paramount to understanding cosmological theories. It is well known that many underlying processes are triggered or enhanced due to galaxy interaction. The clearest result of interaction is the disruption of galactic morphology, the acceleration of disks to spheroidal systems and the formation of tidal features \citep{1972ApJ...178..623T, 1977ApJ...212..616T, 2005MNRAS.357..753G, 2009MNRAS.397..802H}, increases in star formation rates when compared to the general galaxy population \citep{1991ApJ...370L..65B, 2006ApJ...652...56B, 2014MNRAS.437.2137S, 2015ApJ...807L..16K} and, potentially, their rapid quenching \citep{2013MNRAS.430.1901H, 2023RAA....23i5026D}. There is also evidence to suggest that interaction plays a role in the ignition of active galactic nuclei (AGN) \citep{2011MNRAS.418.2043E, 2015ApJ...806..219C, 2023MNRAS.523.4164H}. Therefore, with interaction having such an affect on the galaxies we see - over a long time period - understanding galaxy interaction is fundamental to understanding galaxy evolution.

To fully map out the effect of galaxy interaction, we must understand its impact on galaxies through the entire merger history and dynamical timescale of the interaction. As each observation only provide snapshots of different parts of the dynamical timescale across different merger histories, it is imperative we have large samples to capture the full history. The most obvious place to start with mapping the full dynamical histories is with simulations. Previous works find sudden increases in star formation occurs early in the dynamical timescale of the interaction \citep{2008MNRAS.384..386C, 2019MNRAS.490.2139R}, however, confirming this definitive link observationally remains elusive \citep{2023ApJ...958...96R}. The same studies could also be done for nuclear activation as well as quenching of interacting systems \citep{2011MNRAS.418.2043E, 2018PASJ...70S..37G, 2023ApJ...942..107S}.

Achieving such study in the dynamical history of an interaction with observations is more difficult. While we cannot directly measure the full dynamical timescale for individual systems, we can create large samples of interacting galaxies which are representative over the dynamical history. The first interacting galaxy sample (the \citet{1966ApJS...14....1A} catalogue) contained only 318 systems, while more recent samples have been created either by machine learning classification \citep{2019A&A...626A..49P, 2023A&A...669A.141S}, visual classification by citizen scientists \citep{2010MNRAS.401.1043D} or by photometric parameterisation \citep{2004AJ....128..163L, 2023MNRAS.522....1N}. With larger samples, we can approximate different parts of the dynamical time them into their constituent stages. However, each sample has always been plagued by contamination and a loss of statistical significance when broken down into stages. It is notoriously difficult, from morphology alone, to select interacting or merging galaxies.

Often, the stage of an interaction is approximated by measuring different parameters with the projected separation between two systems. Early works showed that there was significant star formation enhancement (SFE) in galaxy pairs with small separations \citep{2003MNRAS.346.1189L,2008AJ....135.1877E, 2022ApJ...940....4S}. A connection was also found between the projected separation in galaxy pairs and the AGN fraction where, once again, the AGN fraction appears to increase with decreasing projected separation \citep{2009MNRAS.399.2172R, 2011MNRAS.418.2043E, 2017MNRAS.465.2671G, 2021ApJ...909..124S}. However, using projected separation as a proxy for the stage in dynamical time of an interacting history overlooks which part of the dynamical history the interaction is at. For instance if a galaxy pair is found to have small projected separation, without visually confirming the morphology, it is difficult to ascertain if the galaxies are just approaching each other or just passed each other. The differences between these being paramount to understanding the effects of interaction on the galaxies involved.

In this work, we split a large sample of interacting galaxies into four specific stages based from their morphology. Each stage is designed to capture different parts of the dynamical time of an interaction. Each interacting or disturbed galaxy is visually classified into its stage, which ranges from galaxies being close pairs, to overlapping and disturbed, to disturbed and distinct to finally coalescing. This follows the classification methods of developing algorithms for staging of interaction\citep{2019MNRAS.490.5390B,2022ApJ...937...97C}.

We elect to use the newly created interacting galaxy catalogue created by \citet{2023ApJ...948...40O} (hereafter, OR23) which found a sample of 21,926 interacting galaxy systems. This is the largest visually classified interacting galaxy catalogue to date. This sample was created using the new ESA Datalabs\footnote{ESA Datalabs: link} platform and the convolutional neural network machine \texttt{Zoobot} \citep{2022MNRAS.509.3966W, 2023JOSS....8.5312W}. This classified all sources in the Hubble Source Catalogue \citep{2016AJ....151..134W} from the Hubble Space Telescope (HST) archives into interacting or non-interacting. Once this was done, a rigorous process of de-duplication, contamination removal and visual inspection was conducted.

We cross match this catalogue with the Cosmic Evolutionary Survey (COSMOS) survey\footnote{DOI: \href{https://www.ipac.caltech.edu/doi/irsa/10.26131/IRSA178}{10.26131/IRSA178}}. This provides us with ancillary data for a subset of the OR23 catalogue. This ancillary data contains many galactic parameters of interest and we explore how they evolve with our defined stages of interaction. We primarily focus on the evolution of the stellar masses, star formation rates and AGN classification from various COSMOS catalogues. We utilise the photometric redshifts available from COSMOS to confirm a set of interacting galaxies and close pairs to draw on trends with projected separation.

This paper is laid out as follows in Section \ref{data} we explain the OR23 and COSMOS catalogues, and describe the process of catalogue matching especially focused on accurate de-duplication and cross-matching between the different catalogues.  Section \ref{method} describes the methods by which we split the interacting systems into stages, and how we identify the secondary galaxies in each pair. In Sections \ref{results:SF_stage} and \ref{results:AGN_stage} we show our results of star formation and active galactic nuclei evolution with interaction stage and present an initial discussion. These are followed in Section \ref{discussion} where we compare to previous works and put our results in the context of the field. Finally, Section \ref{conclusion} we make concluding remarks and discuss future work to better our constraints.

Where necessary, we use a Flat $\Lambda$CDM cosmology with $H_0$ = $70$\,km/s/Mpc and $\Omega_M = 0.3$. Hereafter in this paper, when referring to an interacting galaxy we are referring to a galaxy which has undergone one or multiple flybys by a secondary galaxy and caused tidal disturbance. A merging galaxy is the final state of these flybys, where two or more systems have or are about to coalesce to form a morphologically irregular system.

\section{DATA: Catalogue Matching \& Secondary Identification} \label{data}
\subsection{The O'Ryan+23 Catalogue}
\noindent The interacting galaxy catalogue we draw from is the OR23 catalogue of interacting galaxies. These systems are classified based on their visual morphology and, while the full process of its creation are detailed there, we will briefly describe it here. The morphological classifications were made by a convolutional neural network (CNN) - \texttt{Zoobot} - which searched all final product data in the \textit{Hubble} Legacy Archive. While a CNN primarily classifies images by extracting features from them \texttt{Zoobot} differs by extracting features from an image and reducing them to generalised representations. By learning the generalise representation of different galaxy morphologies, \texttt{Zoobot} is incredibly flexible and easy to transfer learn on different tasks and datasets. This is further reinforced by training it on citizen scientist responses to the Galaxy Zoo: DECaLS \citep{2022MNRAS.509.3966W} and using it to make classifications on multiple morphology questions at once. In OR23, using a small dataset of citizen science-classified interacting galaxies from Galaxy Zoo \citep{2008MNRAS.389.1179L} \texttt{Zoobot} was finetuned to \textit{Hubble} data.

Cutouts of each source were created as single-band \textit{F814W} images at 150 $\times$ 150 pixels across. Source coordinates were found in the \textit{Hubble} Source Catalogue \citep{2016AJ....151..134W}. This corresponded to just over 92 million cutouts created and classified. Due to well known issues of classifying interacting galaxies by morphology alone (such as close pair contamination) an extremely conservative prediction score from \textit{Zoobot} was required to make a positive classification. This resulted in a very pure, but highly incomplete catalogue of 21,926 interacting galaxy systems. This is the largest catalogue of interacting galaxies directly found to date.

The OR23 catalogue only provided the coordinates and \textit{Hubble} Source catalogue ID for each source. Therefore, we must either calculate our own ancillary data or cross match with existing catalogues containing it. We elect to cross match the OR23 catalogue with the numerous catalogues of the COSMOS survey. The overall survey has numerous sub-catalogues ranging from calculations of galactic physical parameters to predictions on AGN classifications.

\subsection{The COSMOS2020 Catalogue}
In the COSMOS survey, there are multiple catalogues of galactic parameters such as star formation rates (SFR), stellar masses and line emissions. We elect to specifically use the COSMOS2020 catalogue \citep{2022ApJS..258...11W}. This catalogue has a wealth of ancillary information for just over 1.7 million sources in a 2 square degree area of the sky. Each source found in this area has been analysed with well known astronomical software (for our purposes, primarily  LePhare \citep{1999MNRAS.310..540A, 2006A&A...457..841I}, EAzY \citep{2008ApJ...686.1503B} and FAST \citep{2017MNRAS.465.3390A}). These provide estimates of the physical parameters of each source in the COSMOS2020 catalogue based on its measured broadband photometry. For our purposes, the most important parameters these calculate are the stellar masses and the estimated star formation rates of each source. We opt to simply use one of these two measurements in our analysis. For the stellar mass we use the LePhare measurement and for the star formation rates we use the EAzY/FAST measurements. We compare the output masses and SFRs for our sources in Figure \ref{fig:difference-measures}. As shown, there can be wide scatter and disagreement between the two algorithms. We, therefore, elect to only use one software for each parameter and not to match the stellar masses and SFRs between them.

\begin{figure}
    \centering
    \includegraphics[width=0.45\textwidth]{figures/mass-sfr-scatter.pdf}
    \caption{Comparison of the measures of stellar mass and SFR using either LePhare or EAzY/FAST photometric codes to calculate them. If the algorithms agreed perfectly, the sources would lie on the blue 1:1 line. \textit{Left}: The scatter in the stellar masses between softwares. As shown, EAzY often seems to find larger stellar masses when compared to LePhare. \textit{Right}: Scatter in SFRs when measured with LePhare or EAzY. There is a large difference between the measures of LePhare and EAzY here, and therefore we opt to only use EAzY measures of the SFR through this work.}
    \label{fig:difference-measures}
\end{figure}

We cross match between the OR23 catalogue with the COSMOS2020 catalogue by matching coordinates between them. We search the COSMOS2020 catalogue for any sources within 10" of the OR23 coordinates and make a list of any COSMOS2020 sources. If more than one COSMOS2020 source is found within this radius we take the source with the smallest angular separation from the OR23 coordinates as the correct source. Once we have identified one COSMSO2020 source for every OR23 ID, we de-duplicate based on COSMOS2020 ID and redo the coordinate matching process with any duplicate matches. If no further COSMOS2020 sources were within 10" of the OR23 source, then we the OR23 source not in the COSMOS2020 catalogue. We find 3,786 of the OR23 sources exist in the COSMOS2020 catalogue.

Once matched, we remove any sources with non-physical photometric measurements for the stellar mass or star formation rates. We then further reduce our sample by only keeping sources within the a mass range of $6.5 \leq \log_{10} \text{M}_{*}(\text{M}_{\odot}) \leq 12.5$ and a star formation rate range of $-5 < \log_{10} \text{SFR} (M_{\odot}yr^{-1}) \leq 3.5$. To also nullify the effect of surface brightness dimming that may occur when attempting to classify tidal features of interacting galaxies, we institute a redshift cut of $z \leq 1.2$. This also makes our sample match the catalogue that will be later used to investigate the incidence of our sample of interacting galaxies with environment (detailed in Section \ref{data:environ}). Applying these cuts reduces our sample size to 3,689 interacting galaxies.

\subsection{Secondary Identification}\label{sec:sec-ident}
\noindent As the OR23 catalogue only contains source coordinates and IDs, we must also manually identify the secondaries of many of our interacting systems. To find the secondaries, we apply three steps. First, a cutout surrounding each OR23 source was created. These cutouts were from the COSMOS cutout service, selecting HST-ACS tiles in the $F814W$ filter. Each cutout had a 30" radius (corresponding to 1001 $\times$ 1001 pixels). The original OR23 cutout was also displayed next to the enlarged cutout. We annotate each cutout with each sources COSMOS2020 ID and measured photometric redshift and error. By annotating each cutout with the sources photometric redshift and ID, we could visually assessed each cutout and give one of four following classifications to each: system disturbed but secondary could not be identified; secondary could be identified; cannot confirm galaxy is interacting; null redshift; incorrect primary assigned.

To classify a secondary galaxy for our primary one, it had to be within the cutout we were visually assessing and within the recorded error of the primary photometric redshift. Using photometric redshift cutoffs in this way is often done when calculating environment parameters \citep[e.g][]{2006MNRAS.373..469B} or defining interacting galaxies by close pairs \citep[e.g][]{2022ApJ...940....4S}. A null redshift is defined as one outwith our redshift limits or NaN. A minority of the cutouts we visually assessed were found to have the incorrect primary at the centre. In these cases, we record the correct primary galaxy ID and extract the ancillary data from the COSMOS2020 catalogue. We then attempt to find the secondary galaxy again for the corrected primary.

@@

\begin{figure*}
    \centering
    \includegraphics[width=\textwidth]{figures/cutouts_ex.pdf}
    \caption{An example of each visual classification made on the cross matched sample. From left to right these are the secondary could not be identified, the primary had a clear secondary, the primary could not be reliably classified as an interacting galaxy, the redshift was null and the incorrect primary was identified in OR+23. Based on these classifications, we either add the secondary galaxies to the sample or we remove the contamination from it. These images are 30" across using the COSMOS cutout service, selecting HST/ACS tiles as the basis for the observations in the $F814W$ filter.}
    \label{fig:secondary_selection}
\end{figure*}

% Define our classifications above.
While initially surprising that the majority of our systems could not have a secondary identified, we found that it was mostly due to limitations in the COSMOS catalogue or the way in which we conduct our secondary identification. Each potential secondary must have a COSMOS ID associated with it, however, when two systems were very close together and small enough they would be identified under a single COSMOS ID despite being two separate systems. The same also occurred when two systems were merging or interacting. The tidal features connecting the systems or coalescing systems would only be identified under one ID in the catalogue. Figure \ref{fig:secondary_selection} shows an example of two systems being close enough together that they have been identified under a single COSMOS2020 entry. Figure \ref{fig:secondaries_found} shows this disparity with the different types of interaction we observe.

\begin{figure}
    \centering
    \includegraphics[width=0.45\textwidth]{figures/visualisation_classification.pdf}
    \caption{Where a secondary could be identified at different stages in the interaction. The reasoning for such disparity in secondaries identified is due to the relative distance each the secondary would be from the primary at each stage. For a close pair, we often found the secondary galaxy, but a minority of these were so close together that the entire system was given a single COSMOS ID. The same was true for those interacting systems approaching pericentre or merging. When the secondary was near apocentre, often it would be outwith the cutout we were using for visual classification.}
    \label{fig:secondaries_found}
\end{figure}

Those galaxies which were found to be contamination (i.e., could not be reliably classified as interacting or having a null redshift) were removed from our sample. Galaxies that could not be reliably classified as interacting as they showed little tidal distortion or had no neighbouring systems at a matching redshift were also removed. These systems were often overlapping but at different redshifts, or were systems with irregular morphologies of spiral arms or many clumpy galaxies. There was also many systems that were at high redshift ($z \geq 1$) where the resolution of the cutouts meant that features could not be discerned visually.

Our final classification type was that the incorrect galaxy had been identified as the priamry galaxy. This was the case for 149 systems. These were systems where the interacting galaxies were clearly in the cutout but some tidal debris or some nearby system had been cross matched from COSMOS. We reassign these systems to the correct COSMOS IDs and then take them through the secondary identification process.

By the end of this pipeline, we find a sample size of 3,829 interacting galaxies. We conduct a de-duplication based on the COSMOS2020 ID, which reduces the sample back down to 3,547 interacting galaxies. This shows that many of the secondaries we had identified for the different systems were already found in the OR23 catalogue. The remaining systems are each visually confirmed interacting and disturbed galaxies based on their morphology and photometric redshift as measured in the COSMOS2020 catalogue.

For each galaxy in our sample, we find and define a mass- and redshift-matched control galaxy to investigate differences in their galactic parameters. We find these control galaxies from the COSMOS2020 catalogue. All galaxies within 0.01 dex of our sub-samples' stellar mass are selected from the catalogue, and within a $\pm$0.01 redshift slice. We then define the control galaxy as the system furthest from our interacting galaxy. Each control galaxy is then visually confirmed to be non-interacting and have no nearby pairs. We also confirm that it does not already exist in the catalogue, and ensure there is no duplication in the control sample. Figure \ref{fig:matched-distributions} shows the mass distribution of our paired sub-sample of galaxies and their control, showing a mass distribution which is approximately the same as the interacting sample. With the previously defined cuts, we find a control galaxy for all but one of our interacting galaxies.

\begin{figure}
    \centering
    \includegraphics[width=0.45\textwidth]{figures/mass-matching-pairs.pdf}
    \caption{The mass distribution of the paired interacting galaxy sample and the control sample. Both primary and secondary galaxies are within this distribution. From our sample of 4,181 interacting galaxies morphologically identified, 834 are confirmed galaxy pairs. This means that the secondary in the interacting galaxy system has been identified from morphology and photometric redshifts.}
    \label{fig:matched-distributions}
\end{figure}

\subsection{Finding Additional Systems}
\noindent As a result of using visual classification to find the secondary galaxy in each interaction, we were able to also confirm other interacting systems which had not been found in the OR23 catalogue. Primarily, these extra interacting galaxies are from systems which had more than one galaxy involved in the interaction. Our pipeline was built to only find a primary and secondary galaxy and, therefore, we add these extra systems through another pipeline. Other interacting galaxies that were added were low redshift systems which would have appeared very large in the small OR23 cutout. By looking at the larger COSMOS2020 cutouts, we are able to recover these galaxies and add them to this sample.

In total, we found an extra 841 interacting systems that we could add to our sample. Upon conducting a de-duplication of these with the sample already found, this was reduced to 634 interacting systems. This gives us a total sample size of 4,181 which we use through the rest of this work.

>>>>

% Abstract of the paper
\begin{abstract}
   Galaxy interaction and merging have fundamental impacts on galactic evolution. From increases in star formation, to morphology, to activation of nuclear activity, galaxy interaction is thought to play a role. However, at what point in the interaction do these effects occur? In this work, we divide a large sample of interacting galaxies in the COSMOS field into four distinct stages based on their interaction morphology. Each stage captures a different part of the dynamical timescale of interaction. We then probe the relation between stage, mass, star formation, and nuclear activity. We find that the distribution of star formation rates rapidly changes with stage for an equivalent mass distribution. We find a shrinking red sequence with stage, and an increase in the fraction of starburst galaxies from 0.1 as close pairs to 0.38 at the point of coalescence. We also find that the fraction of active galactic nuclei is static through most of the interaction, until it increases at coalescence. We discuss these results, and compare them with the more prevalent stage diagnostic of projected separation between galaxy pairs. We find morphology and stage dependence in the rate of star formation which is not captured by using the projected separation. We show that the point in time of the underlying dynamical timescale an interaction occupies is as important to consider as its projected separation.

\end{abstract}

@@

\section{Introduction}\label{introduction}
\noindent To fully map out the effect of galaxy interaction, we must understand its impact through various merger histories and across the full dynamical timescale of the interaction. The dynamical timescale is the full time of the interaction, from the two galaxies approaching each other to flying by and escaping or merging. We cannot observe a full, continuous interaction and have to rely on many observations to provide snapshots of the dynamical timescale we can piece together. This, however, makes it difficult to conclusively find link the dynamical timescale to physical processes. Often, observational samples are too small to be fully representative or deep studies of individual systems are required to identify the dynamical timescale \citep{2000ApJ...530..660B, 2012A&A...539A..45L, 2014A&A...567A.132W}. We gain insights from works focused on simulations, where the dynamical timescale can be directly recorded. They find sudden increases in star formation occurs at the mid-point of the dynamical timescale, as the two systems flyby \citep{2008MNRAS.384..386C, 2019MNRAS.490.2139R, 2021MNRAS.503.3113M}. However, confirming this observationally remains elusive \citep{2023ApJ...958...96R}. Observational confirmation is also difficult with nuclear activation and quenching of interacting systems \citep{2011MNRAS.418.2043E, 2018PASJ...70S..37G, 2023ApJ...942..107S}.

While cannot directly measure the full dynamical timescale for individual systems, we can create large samples of interacting galaxies contain systems spanning the full dynamical history. Attempts to create such a sample have been either by machine learning classification \citep{2019A&A...626A..49P, 2023A&A...669A.141S}, visual classification by citizen scientists \citep{2010MNRAS.401.1043D} or by photometric parameterisation \citep{2004AJ....128..163L, 2023MNRAS.522....1N}. However, as stated previously, each sample has always been plagued by contamination and a loss of statistical significance when broken down into stages.

An alternate approach to directly measuring the dynamical timescale of an interaction has been to infer it from the projected separation between the two systems. Early works showed that there was significant star formation enhancement (SFE) in galaxy pairs with small separations \citep{2003MNRAS.346.1189L,2008MNRAS.385.1903L, 2008AJ....135.1877E, 2022ApJ...940....4S}. A connection was also found between the projected separation in galaxy pairs and the AGN fraction where, once again, the AGN fraction appears to increase with decreasing projected separation \citep{2009MNRAS.399.2172R, 2011MNRAS.418.2043E, 2017MNRAS.465.2671G, 2021ApJ...909..124S}. However, using projected separation as a proxy for the stage in dynamical time of an interacting history overlooks which part of the dynamical history the interaction is. For instance, if a galaxy pair is found to have small projected separation, without visually confirming the morphology, it is difficult to ascertain if the galaxies are just approaching each other or have just passed each other. Extended morphological tracers can help break this degeneracy, which is critical to our understanding of the effects of interaction.

In this Chapter, we split our large sample of interacting galaxies found in Chapter \ref{chapter2} into four specific stages based on their morphology. Each stage is designed to capture different parts of the dynamical time of an interaction. The stages range from galaxies being close pairs, to overlapping and disturbed, to disturbed and distinct to finally coalescing. This follows the classification methods of developing algorithms for staging of interaction\citep{2019MNRAS.490.5390B,2022ApJ...937...97C}.

We focus here on the subset of our sample in the Cosmic Evolutionary Survey (COSMOS) survey\footnote{DOI: \href{https://www.ipac.caltech.edu/doi/irsa/10.26131/IRSA178}{10.26131/IRSA178}}. This provides us with ancillary data which contains many galactic parameters of interest and we explore how they evolve with our defined stages of interaction. We primarily focus on the evolution of the stellar masses, star formation rates and AGN classification from various COSMOS catalogues. We use the photometric redshifts available from COSMOS to confirm a set of interacting galaxies and close pairs to draw on trends with projected separation.

This Chapter is laid out as follows in Chapter \ref{data} we briefly summarise the COSMOS catalogue, and describe the process of catalogue matching especially focused on accurate de-duplication and cross-matching. Chapter \ref{method} describes the methods by which we split the interacting systems into stages, and how we identify the secondary galaxies in each pair. In Sections \ref{results:SF_stage} and \ref{results:AGN_stage} we show our results of star formation and active galactic nuclei evolution with interaction stage and present an initial discussion. These are followed in Chapter \ref{discussion} where we compare to previous works and put our results in the context of the field. Finally, in Chapter \ref{conclusion} we make concluding remarks and discuss future work to better our constraints.

\section{Data: Catalogue Matching \& Secondary Identification} \label{data}
\subsection{The COSMOS2020 Catalogue}
In the COSMOS survey, there are multiple catalogues of galactic parameters such as star formation rates (SFR), stellar masses and line emissions. We elect to specifically use the COSMOS2020 catalogue \citep{2022ApJS..258...11W}. This catalogue has a wealth of ancillary information approximately 1.7 million sources in a 2 square degree area of the sky. Each source has been analysed with well known astronomical software (for our purposes, primarily \texttt{\texttt{LePhare}} \citep{1999MNRAS.310..540A, 2006A&A...457..841I}, \texttt{\texttt{EAzY}} \citep{2008ApJ...686.1503B} and \texttt{\texttt{FAST}} \citep{2017MNRAS.465.3390A}). These provide estimates of the physical parameters of each source based on its measured broadband photometry. For our purposes, the parameters of interest are the stellar masses and the estimated star formation rates of each source. For the stellar mass we use the best fit \texttt{LePhare} measurement and for the star formation rates we use the best fit \texttt{EAzY}/\texttt{FAST} measurements. We compare the output masses and SFRs for our sources in Figure \ref{fig:difference-measures}. As shown, there can be wide scatter and disagreement between the two algorithms. We tested different combinations of \texttt{EAzY} and \texttt{LePhare} stellar masses and SFRs to find which would be best to use in this work. We found that using the same algorithm in both parameters revealed the degeneracies of using discrete templates to fit the broad band photometry in the underlying parameter space. Therefore, we elect to inter-mix the algorithms and use the best fit \texttt{LePhare} stellar mass and best fit \texttt{EAzY} SFR measurement.

\begin{figure}
\centering
\includegraphics[width=\columnwidth]{figures/mass-sfr-scatter.pdf}
\caption[Comparison of the measures of stellar mass and SFR using either \texttt{LePhare} or \texttt{EAzY}/\texttt{FAST} photometric codes to calculate them.]{Comparison of the measures of stellar mass and SFR using either \texttt{LePhare} or \texttt{EAzY}/\texttt{FAST} photometric codes to calculate them. If the algorithms agreed perfectly, the sources would lie on the blue 1:1 line. \textit{Left}: The scatter in the stellar masses between softwares. As shown, \texttt{EAzY} often seems to find larger stellar masses when compared to \texttt{LePhare}. \textit{Right}: Scatter in SFRs when measured with \texttt{LePhare} or \texttt{EAzY}.}
\label{fig:difference-measures}
\end{figure}

Our sample of interacting galaxies has been de-duplicated, but the COSMOS2020 catalogue is not specifically a de-duplicated merger catalogue. We cross match between our sample and the COSMOS2020 catalogues using a position search within $10^{\prime\prime}$ of our sample coordinates. Once we have identified the nearest COSMOS2020 source for each interacting galaxy ID, we de-duplicate based on COSMOS2020 ID and redo the coordinate matching process with any duplicate matches. If no further COSMOS2020 sources were within 10" of the source, then we classify the source as not in the COSMOS2020 catalogue. We find 3,786 of the our sources exist in the COSMOS2020 catalogue.

Once matched, we remove any sources with non-physical photometric measurements for the stellar mass or star formation rates. We then further reduce our sample by only keeping sources within a mass range of $6.5 \leq \log_{10} \text{M}_{*}(\text{M}_{\odot}) \leq 12.5$ and a star formation rate range of $-5 < \log_{10} \text{SFR} (\text{M}_{\odot}\text{yr}^{-1}) \leq 3.5$. We also opt to institute a redshift cut of $z \leq 1.2$. Beyond this redshift, we find that identification of tidal features becomes difficult due to surface brightness dimming and we risk mis-identifying the stages of the interacting galaxies. This also matches the redshift cut applied to the environment catalogue we cross match with in Chapter \ref{data:environ}. Applying these cuts reduces our sample size to 3,689 interacting galaxies.

\subsection{Secondary Identification}\label{sec:sec-ident}
\noindent As the catalogue described in Chapter \ref{chapter2} only contains source coordinates and IDs, we must also manually identify the secondaries of many of our interacting systems. To find the secondaries, we apply three steps. First, a cutout surrounding each source was created. These cutouts were from the COSMOS cutout service, selecting HST-ACS tiles in the $F814W$ filter. Each cutout had a 30" radius (corresponding to 1001 $\times$ 1001 pixels). The original cutout from Chapter \ref{chapter2} was also displayed next to the enlarged cutout. We annotate each cutout with each sources COSMOS2020 ID and measured photometric redshift and error. By annotating each cutout with the sources photometric redshift and ID, we visually assessed each cutout and gave one of the four following classifications to each: system disturbed but secondary could not be identified; secondary could be identified; cannot confirm galaxy is interacting; null redshift (0 or NaN); incorrect primary assigned.

To associate a secondary galaxy for each primary, the galaxy had to be within the cutout we were visually assessing and within the recorded error of the primary photometric redshift. Using photometric redshift cutoffs in this way is often done when calculating environment parameters \citep[e.g][]{2006MNRAS.373..469B} or defining interacting galaxies by close pairs \citep[e.g][]{2022ApJ...940....4S}. A null redshift is defined as outwith our redshift limits, 0 or NaN. A minority of the cutouts we visually assessed were found to have the incorrect primary at the centre. In these cases, we record the correct primary galaxy ID and extract the ancillary data from the COSMOS2020 catalogue. We then attempted to find the secondary galaxy again for the corrected primary.

@@

\begin{figure*}
\centering
\includegraphics[width=\textwidth]{figures/cutouts_ex.pdf}
\caption[An example of each visual classification made on the cross matched sample.]{An example of each visual classification made on the cross matched sample. These are: (A) where the secondary could not be identified, (B) the primary had a clear secondary, (C) the primary could not be reliably classified as an interacting galaxy, (D) the redshift was null and (E) the incorrect primary was identified. Based on these classifications, we either add the secondary galaxies to the sample or we remove the contamination from it. These images are 30" across using the COSMOS cutout service, selecting HST/ACS tiles as the basis for the observations in the $F814W$ filter.}
\label{fig:secondary_selection}
\end{figure*}

% Define our classifications above.
While initially surprising that the majority of our systems could not have a secondary identified, we found that it was mostly due to limitations in the COSMOS catalogue or the way in which we conduct our secondary identification. Each potential secondary must have a COSMOS ID associated with it, however, when two systems are very close together and small enough they were identified under a single COSMOS ID despite being two separate systems. The same also occurred when two systems were merging or interacting. The tidal features connecting the systems or coalescing systems would only be identified under one ID in the catalogue. Figure \ref{fig:secondary_selection} panel C) shows an example of two systems being close enough together that they have been identified under a single COSMOS2020 entry. Figure \ref{fig:secondaries_found} shows this disparity with the different types of interaction we observe.

\begin{figure}
\centering
\includegraphics[width=\columnwidth]{figures/visualisation_classification.pdf}
\caption[Where a secondary could be identified at different stages in the interaction.]{Where a secondary could be identified at different stages in the interaction. The reasoning for such disparity in secondaries identified is due to the relative distance each the secondary would be from the primary at each stage. For a close pair, we often found the secondary galaxy, but a minority of these were so close together that the entire system was given a single COSMOS ID. The same was true for those interacting systems approaching pericentre or merging. When the secondary was near apocentre, often it would be outside the cutout we were using for visual classification.}
\label{fig:secondaries_found}
\end{figure}

Those galaxies which were found to be contamination (i.e., could not be reliably classified as interacting as they showed little tidal distortion or had no neighbouring systems at a matching redshift or having a redshift of 0) were removed from our sample. Galaxies that could not be reliably classified as interacting were also removed. These systems were often overlapping but at different redshifts, or were systems with irregular morphologies of spiral arms or a clumps were present. There was also many systems that were at high redshift ($z > 1$) where the resolution of the cutouts meant that features could not be discerned visually.

Our final classification type was that the incorrect galaxy had been identified as the primary galaxy. This was the case for 149 systems. These were systems where the interacting galaxies were clearly in the cutout but some tidal debris or some nearby system had been matched from COSMOS. We reassign these systems to the correct COSMOS IDs and then take them through the secondary identification process.

By the end of this selection method, we find a sample size of 3,829 interacting galaxies. We conduct a de-duplication based on the COSMOS2020 ID, which reduces the sample back down to 3,547 interacting galaxies. The remaining systems are each visually confirmed interacting and disturbed galaxies based on their morphology and photometric redshift as measured in the COSMOS2020 catalogue.

For each galaxy in our sample, we define a mass- and redshift-matched control galaxy to investigate differences in their galactic parameters. We find these control galaxies from the COSMOS2020 catalogue. All galaxies within 0.01 dex of our samples stellar mass are selected from the catalogue, and within a $\pm0.01$ redshift slice. We then define the control galaxy as the system furthest from our interacting galaxy. Each control galaxy is then visually confirmed to be non-interacting and have no nearby pairs. We also confirm that it does not already exist in the catalogue, and ensure there is no duplication in the control sample. Figure \ref{fig:matched-distributions} shows the mass distribution of our paired sub-sample of galaxies and their control, showing a mass distribution which is approximately the same as the interacting sample. With the previously defined cuts, we find a control galaxy for all but one of our interacting galaxies.

\begin{figure}
\centering
\includegraphics[width=\columnwidth]{figures/mass-matching-pairs.pdf}
\caption[The mass distribution of the paired interacting galaxy sample and the control sample.]{The mass distribution of the paired interacting galaxy sample and the control sample. Both primary and secondary galaxies are within this distribution.}
\label{fig:matched-distributions}
\end{figure}

\subsection{Finding Additional Systems}
\noindent As a result of using visual classification to find the secondary galaxy in each interaction, we were able to also confirm other interacting systems which had not been found in our catalogue. Primarily, these extra interacting galaxies are from systems which had more than two galaxies involved in the interaction. Our selection process was built to only find a primary and secondary galaxy and, therefore, we add these extra systems into our sample manually. Other interacting galaxies that were added were low redshift systems which would have appeared to completely fill the cutout of the classification process in Chapter \ref{chapter2}. By looking at the larger COSMOS2020 cutouts, we are able to recover these galaxies and add them to this sample.

In total, we found an extra 841 interacting systems that we could add to our sample. Upon conducting a de-duplication of these with the sample already found, this was reduced to 634 interacting systems. This gives us a total flux-limited sample size of 4,181.
